---
title: "Fragmentation matching Round 2"
author: "Jessica Trowbridge"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
library(stringr)

```

## Fragmentation data 
## POSITIVE MODE
#try merging with the summary data + SSI DB data instead of the filtered list? 

The following code from Vincent gives one line per consensus (when there are multiple matches to ID. They are separated by a comma which does not allow us to merge the dataset from Junesoo)

```{r}
###Upload results data from GNPS MS/MS annotation (WWBC database)
node <- read.csv("~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation matching/node_attributes_table_pos1.tsv", header = T, sep = "\t") #Postive Mode
node$cluster.index<-as.factor(node$cluster.index)

##Unlist chemicals ID
node$ConsensusID<-as.character(node$ConsensusID)
node$ConsensusScore<-as.character(node$ConsensusScore)
node$ConsensusSMILES<-as.character(node$ConsensusSMILES)

#this code is to transform the shape of the dataset to have one line per consensus ID  
node1<- node %>%   
  transform(ConsensusID=strsplit(ConsensusID,","),
            ConsensusScore=strsplit(ConsensusScore,","),
            ConsensusSMILES=strsplit(ConsensusSMILES,",")) %>%
  unnest(c(ConsensusID, ConsensusScore, ConsensusSMILES))

#in addition this line of code keeps only those with a consensus ID. it removes fragmentation peaks that do not match. 

node1<-node1[,c(1,2,4,5,17:19)]  

#interested in: cluster.index
# -parent.mass (m/z value)
# -RT mean
# -ConsensusScore (how close is the match between experimental MS/MS and in silico MS/MS)
# -ConsensusSMILES
# -ConsensusID (DTXSID of the chemical)
```

## make unique ID with Mass and RT (using 4 sig digits)

```{r}
node1$mz<-round(node1$parent.mass,1) #parent.mass has 3 sig figs
```

#Then merge with the list of chemicals that we sent JuneSoo. (this includes both Nurse and FF chemicals for fragmentation)
## Import list sent to JuneSoo

```{r}
#data sent for fragmentation. Try Alex's list first then try "by hand"
suspect_list <- read.csv("~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation list to DTSC/POS_filter_features_3MW_30deltaRT_NurseFFCombined.csv",stringsAsFactors = F, header = T) #Postive Mode

# suspect_list <- suspect_list%>%
#   select(c("featureid", "DTXSID","Compound", "Formula", "population"))


#change number of sig figs to match node1 #use MW variable in Alex's combined dataset
suspect_list$mz3<-round(suspect_list$MW,3) #parent.mass has 3 sig figs

#make a new feature ID variable with MW and RT of fragmentation data
suspect_list$MW<-round(suspect_list$MW,1)

suspect_list$RT <- round(suspect_list$RT, 1)
suspect_list$featureid<-as.character(paste(suspect_list$mz3,suspect_list$RT,sep = "-"))

# need to stack the DB_name variables. Have a longer DF. 

suspect_list <- suspect_list %>% 
    mutate(DB_name_new = strsplit(as.character(DB_name), ";")) %>% 
    unnest(DB_name_new)

```

## merge lists 

```{r}
match_mz <- merge(node1, suspect_list, by.x = "mz", by.y = "MW", all.x = TRUE)

#when include RT, no_matches, the differences are slight, but they don't match up 

# match_feature_id <- merge(node1, suspect_list, by= "featureid") 

```


## Flag matched and non-matched by DTXSID
```{r}
#DB_name = DTXSID
#select the first DB_name of a string to identify confirmed match or not. 


match_mz <- match_mz%>%
  mutate(confirmed_match = ifelse(DB_name_new==ConsensusID, 1, ifelse(is.na(DB_name), 3,0)),
         new_chemical_nomatch = ifelse( is.na(DB_name), 1, 0), 
         unknown_match = ifelse(is.na(DB_name) == "unknown", 1, 0))


#detect those with only C+H (not likely to be read with LC)
# str_extract(match_mz$Formula,"^C\\d+H\\d+$")

match_mz <- match_mz %>%
  mutate(CH_only = ifelse(str_detect(match_mz$formula_cleaned,"^C\\d+H\\d+$") == TRUE, 1,0))

table(match_mz$confirmed_match) #0 is no match, 1 confirmed match N = 32, 

table(match_mz$new_chemical_nomatch) #527 with new potential chemical 

tapply(factor(match_mz$confirmed_match), match_mz$ESI, table) #ESI = population

table(match_mz$unknown_match) #30 chemicals were newly identified


tapply(factor(match_mz$unknown_match), match_mz$ESI, table)



unique_mzrt_pos <- match_mz %>%
  group_by(mz)%>%
  summarise(count=n_distinct(ConsensusID))

# write.csv(unique_mzrt_neg, "2020_unique_mzrt_neg.csv")


# write.csv(unique_mzrt, "2020_unique_mzrt_pos.csv")


# write.csv(match_mz, "2020_12_04selected_features_and_matched_fragmentation_POS.csv")
```

#Merge the fragmentation data with the summary statistics calculated previously
```{r}


#merge with features identified by fragmentation by featureid. 
# pos_candidates$mz<-round(pos_candidates$mz,3) #p
pos_candidates <- read.csv(file = "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation list to DTSC/raw data for selection/20210716POS_features_FF_limited.csv", stringsAsFactors = FALSE)

# pos_candidates <- pos_candidates%>%
#   select(2:50)

pos_candidates$mz<-round(pos_candidates$mz,3)
pos_candidates$rt <- round(pos_candidates$rt, 1)
pos_candidates$featureid<-as.character(paste(pos_candidates$mz,pos_candidates$rt,sep = "-"))


# pos_candidates<-pos_candidates[!duplicated(pos_candidates$featureid),] #maybe I don't want to remove dupes if they have multiple DTXSIDs


# pos_candidates<-pos_candidates[!duplicated(pos_candidates$featureid),] #272 features




merged_features <- merge(match_mz, pos_candidates, by = c("featureid"), all.x = TRUE)

# tox_library <- read.csv(file= "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/Code/data_files_csv/WWBC_MS_database_6.30.21_prepForAnotation - WWBC_MS_database_6.30.21_prepForAnotation.csv", stringsAsFactors = FALSE)
# 
# tox_library <- tox_library[ , 2:39]
# 
# merged_features <- merge(merged_features, tox_library, by.x = "ConsensusID", by.y = "DB_name", all.x = TRUE)
# 
# merged_features<-merged_features[!duplicated(merged_features$ConsensusID),] #21 features. #1 ! confirmed through fragmentation in the positive mode for FF. 

#all above 0.8
# merged_features <- merged_features %>%
#   filter(ConsensusScore >0.5) 


#NURSE DATA MATCH:
Nurse_pos_candidates <- read.csv(file = "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation list to DTSC/raw data for selection/20210727List_matched_filtered_for_Frag_Nurse_POS.csv", stringsAsFactors = FALSE)

Nurse_pos_candidates$mz<-round(Nurse_pos_candidates$mz,3)
Nurse_pos_candidates$rt <- round(Nurse_pos_candidates$time, 1)
Nurse_pos_candidates$featureid<-as.character(paste(Nurse_pos_candidates$mz,Nurse_pos_candidates$rt,sep = "-"))


# Nurse_pos_candidates<-Nursepos_candidates[!duplicated(pos_candidates$featureid),] #maybe I don't want to remove dupes if they have multiple DTXSIDs


# pos_candidates<-pos_candidates[!duplicated(pos_candidates$featureid),] #272 features




NURSE_merged_features <- merge(match_mz, Nurse_pos_candidates, by = c("featureid"), all.x = TRUE)



# #filter out the nurses.
# merged_features <- 
#   merged_features %>%
#   filter(population!="nurses")
# pos_candidates<-pos_candidates[!duplicated(pos_candidates$featureid),] 

#81 firefighter features with unique ConsensusID

# #do with lim_pos that has summary stats and tox data. 
# lim_pos <- read.csv(file = "/Users/trowbridgej/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/Code/20210512Pos_features_FF_limited.csv")
# 
# merged_features_lim <- merge(match_mz, lim_pos, by.x = "ConsensusID", by.y = "DTXSID")

write.csv(merged_features, file = "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation matching/20211007_FFConfirmedfeaturesfrag_pos.csv")
write.csv(NURSE_merged_features, file = "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation matching/20211007_NURSEConfirmedfeaturesfrag_pos.csv")
```

# how many have a df below 90%
```{r}
table(merged_features$df_v1<90) #none have DF <50%
  

```


# Criteria: TO UPDATE 10/07
```{r, eval = FALSE}

 merged_features<- merged_features%>%
 mutate(rt_difference = abs(time-RTMean),
        time_dif_flag <- ifelse(rt_difference <40, 1, 0), #filter to keep 1's
        absMED12 = abs(med_diff1v2), 
        absDF12 = abs(df1v2_difference),
        absMED13 = abs(med_diff1v3),
        absDF13 = abs(df1v3_difference),
        sigdif1v2 = ifelse(conc_p_V2 < 0.1, 1, 0), 
        sigdif1v3 = ifelse(conc_p < 0.1, 1, 0),
        df_dif1v2 = ifelse(absDF12 !=0, 1, 0), 
        df_dif1v3 = ifelse(absDF13 != 0, 1,0))
 



merged_features$bc_chem_flag <- round(rowSums(merged_features[ , c("E2Up_bin", "ERactive_bin", "P4Up_bin", "pesticidemammarytumors_bin", "MC")], na.rm = TRUE),0)


merged_features <- merged_features%>%
  mutate(PFAS1 = ifelse(PFAS >=1, 1, NA))
merged_features$FRs <- as.numeric(merged_features$FRs)

merged_features$chem_flag <- rowSums(merged_features[ , c("nitroPAH_bin", "PFAS1",  "FRs")], na.rm = TRUE)                                        

merged_features  <- merged_features %>%
  mutate(has_fun= ifelse(str_detect(Formula.x,"F") == TRUE, "1", ifelse(str_detect(Formula.x,"Cl") == TRUE,"1", ifelse(str_detect(Formula.x,"Br") == TRUE, "1", 0))),
         fun_group = ifelse(str_detect(Formula.x,"F") == TRUE, "F", ifelse(str_detect(Formula.x,"Cl") == TRUE,"Cl", ifelse(str_detect(Formula.x,"Br") == TRUE, "Br", 0)))
        )

# merged_features <- merged_features %>%
#   mutate(sigdif= ifelse(sigdif1v2==1|sigdif1v3==1,1,0), 
#          DF_dif = ifelse(df_dif1v2 ==1|sigdif1v3==1,1,0)
#          ) 

merged_features$flag_sums <- rowSums(merged_features[ ,c("sigdif1v2", "sigdif1v3","df_dif1v2","df_dif1v3","chem_flag", "chem_flag", "bc_chem_flag")], na.rm = TRUE)   


merged_features <- merged_features %>%
  mutate(MSMS_name = Compound.y)


#merge Pos list with OPERA data downloaded from EPA tox

tox_d_pos <- read.csv("data for analysis/Comptox_OPERA_frag_POS.csv")


merged_features <- merge(merged_features, tox_d_pos, by.x = "ConsensusID", by.y = "DTXSID")
# write.csv(merged_features, file = "FF_framented_features_with_toxdata_stats2021_02_18POS.csv")

```



# NEGATIVE MODE



```{r}
###Upload results data from GNPS MS/MS annotation (WWBC database)
node_neg <- read.csv("~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation matching/node_attributes_table_neg.tsv", header = T, sep = "\t") #Negative Mode
node_neg$cluster.index<-as.factor(node_neg$cluster.index)

##Unlist chemicals ID
node_neg$ConsensusID<-as.character(node_neg$ConsensusID)
node_neg$ConsensusScore<-as.character(node_neg$ConsensusScore)
node_neg$ConsensusSMILES<-as.character(node_neg$ConsensusSMILES)



#this code is to transform the shape of the dataset to have one line per consensus ID  
node_neg1<- node_neg %>%   
  transform(ConsensusID=strsplit(ConsensusID,","),
            ConsensusScore=strsplit(ConsensusScore,","),
            ConsensusSMILES=strsplit(ConsensusSMILES,",")) %>%
  unnest(c(ConsensusID, ConsensusScore, ConsensusSMILES))

#in addition this line of code keeps only those with a consensus ID. it removes fragmentation peaks that do not match. 

node_neg1<-node_neg1[,c(1,2,4,5,17:19)]  

#interested in: cluster.index
# -parent.mass (m/z value)
# -RT mean
# -ConsensusScore (how close is the match between experimental MS/MS and in silico MS/MS)
# -ConsensusSMILES
# -ConsensusID (DTXSID of the chemical)
```

## make unique ID with Mass and RT (using 1 sig digits) 
more digits and there is no matching

```{r}
node_neg1$mz3<-abs(round(node_neg1$parent.mass,3)) #parent.mass has 3 sig figs
node_neg1$mz<-abs(round(node_neg1$parent.mass,1)) #parent.mass has 3 sig figs

```

#Merge with the list of chemicals that we sent JuneSoo. (this includes both Nurse and FF chemicals for fragmentation)
## Import list sent to JuneSoo

```{r}
suspect_list_neg <- read.csv("~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation list to DTSC/NEG_filter_features_3MW_30deltaRT_NurseFFCombined.csv", header = T) #Negative

#change number of sig figs to match node1
suspect_list_neg$MW3 <- round(suspect_list_neg$MW, 3)
suspect_list_neg$MW<-round(suspect_list_neg$MW,1)
suspect_list_neg$RT <- round(suspect_list_neg$RT, 1)
suspect_list_neg$featureid<-as.character(paste(suspect_list_neg$MW3,suspect_list_neg$RT,sep = "-"))


# need to stack the DB_name variables. Have a longer DF. 

suspect_list_neg <- suspect_list_neg %>% 
    mutate(DB_name_new = strsplit(as.character(DB_name), ";")) %>% 
    unnest(DB_name_new)


#combine temp_parent_id with the ids from dat and add to test_df



```

## merge lists 

```{r}
head(suspect_list_neg$MW)
head(node_neg1$mz)

table(suspect_list_neg$MW %in% node_neg1$mz)

match_mz_neg <- merge(node_neg1, suspect_list_neg, by.x = "mz", by.y = "MW")

#when include RT, no_matches, the differences are slight, but they don't match up 

# match_feature_id <- merge(node1, suspect_list, by= "featureid") 

```


## Flag matched and non-matched by DTXSID
```{r}
match_mz_neg <- match_mz_neg%>%
  mutate(confirmed_match = ifelse(DB_name_new==ConsensusID, 1, ifelse(is.na(DB_name), 3,0)),
         new_chemical_nomatch = ifelse( is.na(DB_name), 1, 0), 
         unknown_match = ifelse(is.na(DB_name) == "unknown", 1, 0))




#select those that only have C+H
# str_extract(match_mz_neg$Formula,"^C\\d+H\\d+$")

match_mz_neg <- match_mz_neg %>%
  mutate(CH_only = ifelse(str_detect(match_mz_neg$formula_cleaned,"^C\\d+H\\d+$") == TRUE, 1,0))

table(match_mz_neg$confirmed_match) # 0 is no match, 1 is confirmed match (n=85). 2 is new chemical. 

tapply(factor(match_mz_neg$confirmed_match), match_mz_neg$ESI, table)


# write.csv(match_mz_neg, "2020_12_04selected_features_and_matched_fragmentation_NEG.csv")
```


```{r}


unique_mzrt_neg <- match_mz_neg %>%
  group_by(mz, confirmed_match)%>%
  summarise(count=n_distinct(ConsensusID))

# write.csv(unique_mzrt_neg, "2020_unique_mzrt_neg.csv")

```

```{r, eval = FALSE}
# #do with lim_pos that has summary stats and tox data. 
# lim_neg <- read.csv(file = "/Users/trowbridgej/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/Code/20210511Neg_features_FF_limited.csv")
# 
# merged_features_limneg <- merge(match_mz_neg, lim_neg, by.x = "ConsensusID", by.y = "DTXSID")
# 
# write.csv(merged_features_limneg, file = "/Users/trowbridgej/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/ESI_neg/20210513_confirmedfeaturesfrag_neg.csv")

```

# Merge the fragmentation data with the summary statistics calculated previously
```{r, eval = FALSE}
# 
# neg_candidates <- read.csv(file = "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/2021Fragmentation list to DTSC/raw data for selection/20210716Neg_features_FF_limited.csv", stringsAsFactors = FALSE)
# #merge with features identified by fragmentation by featureid. 
# 
# 
# neg_candidates<-neg_candidates[!duplicated(neg_candidates$featureid),] 
# 
# #merge
# merged_features_neg <- merge(match_mz_neg, neg_candidates, by = "featureid", all.x = TRUE)

```

# Merge with Tox library loaded when ran code in positive mode

```{r}
tox_library <- read.csv(file= "~/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/Code/data_files_csv/WWBC_MS_database_6.30.21_prepForAnotation - WWBC_MS_database_6.30.21_prepForAnotation.csv")
# 
# 
# tox_library <- tox_library[ , 2:31]

merged_features_test <- merge(match_mz_neg, tox_library, by.x = "ConsensusID", by.y = "DTXSID", all.x = TRUE) #use names for Compound.y

merged_features_neg<-merged_features_neg[!duplicated(merged_features_neg$ConsensusID),] #126 features


merged_features_neg <- merged_features_neg %>%
  filter(ConsensusScore >0.5) #63 features

#filter out the nurses.
merged_features_neg <- merged_features_neg %>%
  filter(population!="nurses")
# write.csv(merged_features_neg, file = "NEGframented_features_with_summary_stats2020_02.csv")
```

#apply selection criteria
```{r}

merged_features_neg <- merged_features_neg%>%
 mutate(RT_difference = abs(time - RTMean), 
        RT_flag = ifelse(RT_difference >40, 0,1), #filter to keep 1's
        absMED12 = abs(med_diff1v2), 
         absDF12 = abs(df1v2_difference),
        absMED13 = abs(med_diff1v3),
        absDF13 = abs(df1v3_difference),
        sigdif1v2 = ifelse(conc_p_V2 <= 0.1, 1, 0), 
        sigdif1v3 = ifelse(conc_p <= 0.1, 1, 0), 
        df_dif1v2 = ifelse(absDF12 !=0, 1, 0), 
        df_dif1v3 = ifelse(absDF13 != 0, 1,0))
 


merged_features_neg$bc_chem_flag <- rowSums(merged_features_neg[ , c("E2Up_bin", "ERactive_bin", "P4Up_bin", "pesticidemammarytumors_bin", "MC")], na.rm = TRUE)

merged_features_neg <- merged_features_neg%>%
  mutate(PFAS = ifelse(PFAS ==1, 1, NA))

# merged_features_neg$PFAS <- as.numeric(merged_features_neg$PFAS)
merged_features_neg$FRs <- as.numeric(merged_features_neg$FRs)

merged_features_neg$chem_flag <- rowSums(merged_features_neg[ , c("nitroPAH_bin", "PFAS",  "FRs")], na.rm = TRUE)                                        
       
merged_features_neg  <- merged_features_neg %>%
  mutate(has_fun= ifelse(str_detect(Formula.x,"F") == TRUE, "1", ifelse(str_detect(Formula.x,"Cl") == TRUE,"1", ifelse(str_detect(Formula.x,"Br") == TRUE, "1", 0))),
         fun_group = ifelse(str_detect(Formula.x,"F") == TRUE, "F", ifelse(str_detect(Formula.x,"Cl") == TRUE,"Cl", ifelse(str_detect(Formula.x,"Br") == TRUE, "Br", 0)))
        )


merged_features_neg <- merged_features_neg %>%
  mutate(sigdif= ifelse(sigdif1v2==1|sigdif1v3==1,1,0))

merged_features_neg$flag_sums <- rowSums(merged_features_neg[ ,c("sigdif1v2", "sigdif1v3", "df_dif1v2","df_dif1v3", "chem_flag", "bc_chem_flag")], na.rm = TRUE)

merged_features_neg <- merged_features_neg %>%
  mutate(MSMS_name = Compound.y)
#merge Pos list with OPERA data downloaded from EPA tox

tox_d_neg <- read.csv("data for analysis/Comptox_OPERA_frag_neg.csv")


merged_features_neg <- merge(merged_features_neg, tox_d_neg, by.x = "ConsensusID", by.y = "DTXSID")

# write.csv(merged_features_neg, file =  "FF_framented_features_with_summary_stats2021_02_16_NEG.csv")

#do with lim_pos that has summary stats and tox data. 
lim_pos <- read.csv(file = "/Users/trowbridgej/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/Code/20210512Pos_features_FF_limited.csv")

merged_features_lim <- merge(match_mz, lim_pos, by.x = "ConsensusID", by.y = "DTXSID")

write.csv(merged_features_lim, file = "/Users/trowbridgej/Box/SHE lab/WWBC/Data analysis/Non-targeted analysis/FF_non-targeted/ESI_pos/20210513_confirmedfeaturesfrag_pos.csv")

```

```{r, eval=FALSE}
# neg_feat_summ <- read.csv(file = "NEGframented_features_with_summary_stats1.csv", stringsAsFactors = FALSE)
# 
# pos_feat_summ <- read.csv(file= "POSframented_features_with_summary_stats1.csv", stringsAsFactors = FALSE)




```

